% Auto-generated LaTeX wrapper for the main Context Zero paper.
% Build with:
%   latexmk -xelatex -shell-escape papers/context_as_objects.tex
\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{fontspec}
\usepackage{microtype}
\usepackage{xcolor}
\usepackage{hyperref}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{enumitem}
\usepackage{setspace}
\usepackage{fancyhdr}
\usepackage{bookmark}
\usepackage[pipeTables,hashEnumerators]{markdown}

\hypersetup{
  colorlinks=true,
  linkcolor=blue,
  urlcolor=blue,
  citecolor=blue,
  pdftitle={Context as Objects: Capsule-Native Coordination for Multi-Agent Teams},
  pdfauthor={Context Zero / CoordPy Programme Draft}
}

\setstretch{1.05}
\setlength{\parskip}{0.55em}
\setlength{\parindent}{0pt}
\setlength{\emergencystretch}{3em}
\setlist[itemize]{topsep=0.35em,itemsep=0.15em,leftmargin=1.5em}
\setlist[enumerate]{topsep=0.35em,itemsep=0.15em,leftmargin=1.7em}

\pagestyle{fancy}
\fancyhf{}
\fancyhead[L]{Context as Objects}
\fancyhead[R]{Context Zero / CoordPy}
\fancyfoot[C]{\thepage}
\setlength{\headheight}{14pt}

\title{\textbf{Context as Objects: Capsule-Native Coordination for Multi-Agent Teams}\\
\large Main paper draft for the Context Zero programme}
\author{Context Zero / CoordPy Programme Draft}
\date{Updated through SDK v3.18\\2026-04-27}

\begin{document}
\maketitle

\begin{abstract}
Multi-agent LLM systems usually treat context as text: prompts, JSON
records, message logs, tool traces, and free-form summaries passed
between roles. That design conflates several distinct problems:
preserving ambiguity, normalizing producer drift, admitting evidence
under budget, decoding evidence jointly, carrying information across
rounds, and deciding which evidence a downstream model should spend
its limited context window on. We argue that this is the wrong
abstraction. The unit of context should be a typed, immutable,
lifecycle-bounded object with explicit budget and provenance.

We implement this view in CoordPy, a capsule-native runtime and
research harness produced by the Context Zero programme. A
capsule is a content-addressed object satisfying six invariants:
identity, type, lifecycle, budget, provenance, and immutability. The
runtime first makes capsules load-bearing in execution rather than
merely post-hoc metadata: capsules are sealed from run setup through
the LLM byte boundary, artifacts are content-addressed at write time,
and lifecycles are mechanically audited. We then extend the same
contract to multi-agent coordination, where agents exchange typed
handoff capsules rather than raw messages.

The paper's main scientific contribution is a decomposition of the
multi-agent context problem into eight coupled structural axes:
(1) admission under budget, (2) intra-round bundle decoding,
(3) cross-round bundle decoding, (4) fixed-vocabulary normalization
of producer drift, (5) layered open-world normalization,
(6) producer-side ambiguity preservation, (7) decoder-side context
packing under bounded token budgets, and (8) end-to-end
producer-plus-decoder composition. Across SDK v3.8-v3.18 we build a
progressive benchmark ladder, R-54 through R-64, that isolates these
axes one by one. Admission alone wins in some regimes but has a named
ceiling. Bundle-aware decoding crosses that ceiling. Cross-round
decoding crosses a further temporal ceiling. Fixed-table
normalization closes a bounded producer-drift gap; layered
normalization closes a wider but still finite open-world gap.
Structured producer protocols close the first real-LLM upstream
ambiguity-erasure gap. Attention-aware capsule context packing closes
a downstream bounded-context gap. End-to-end composition then yields
the first fresh live real-LLM strict +1.000 advance over the
strongest non-composed baseline in the programme.

The strongest positive result in the paper is SDK v3.18's W17
family: on a fresh live \texttt{qwen2.5:14b-32k} probe, a magnitude-hinted
structured producer protocol plus attention-aware capsule packing
yields \texttt{accuracy\_full = 1.000}, while both substrate/FIFO and the
strongest non-composed baseline remain at 0.000 on the same live
benchmark. The same producer-side intervention transfers partially
across model class to a fresh live \texttt{qwen3.5:35b} MoE probe,
preserving the benchmark property 8/8 and yielding a +0.750 strict
gain. The strongest negative result is equally important: SDK v3.18
also proves the programme's first explicit symmetric-corroboration
limit theorem. When gold and decoy are symmetrically corroborated
under the current feature surface, every current closed-form
capsule-native method in the SDK ties FIFO at 0.000 by construction.
This names the next wall sharply rather than leaving it rhetorical.

The paper does not claim that multi-agent context is solved in an
unqualified sense. The strongest honest claim is narrower and more
useful: multi-agent context becomes tractable when evidence is
represented as typed objects and when the runtime explicitly
separates producer-side ambiguity preservation, normalization,
admission, intra-round decoding, cross-round decoding, and
decoder-side bounded-context packing. The next open problem is no
longer "can capsules help?" It is "what richer semantics or learned
disambiguators are required once corroboration, magnitude, round
structure, normalization, and bounded packing are all exhausted?"
\end{abstract}

\tableofcontents
\newpage

\markdownInput{papers/context_as_objects_body.md}

\end{document}
